Draft final report\\ Measles risk assessment, modelling and benefit--cost analysis\\ \vspace{2 mm} {\large David T S Hayman, Tim Carpenter,\\ Jonathan C Marshall, Mick Roberts, Nigel P French}

### 2014-10-30 draftfinalreport.Rnw
major text revisions
minor editing still required. Bulk done

\usepackage{times}
\usepackage{hyperref}
\usepackage{tikz}
%\usepackage{bibunits}
\usepackage{natbib}
\usepackage{graphics}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{indentfirst}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{eurosym}
\usepackage{todonotes}
\usepackage{pdflscape}
\usepackage{booktabs}
\usepackage{array}
\usepackage{rotating}
\usepackage{threeparttable}
\usepackage{blindtext}
\usepackage{dcolumn}
\usepackage{tabularx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{caption2}
\usepackage{caption}
\usepackage{fancyhdr}
\usepackage{placeins}
\usepackage{framed}
\newcommand{\vecA}{\mathbf{A}}
\newcommand{\vecB}{\mathbf{B}}
\newcommand{\vecC}{\mathbf{C}}
\newcommand{\vecD}{\mathbf{D}}
\newcommand{\vecE}{\mathbf{E}}
\newcommand{\vecF}{\mathbf{F}}
\newcommand{\vecH}{\mathbf{H}}
\newcommand{\vecI}{\mathbf{I}}
\newcommand{\vecJ}{\mathbf{J}}
\newcommand{\vecK}{\mathbf{K}}
\newcommand{\vecM}{\mathbf{M}}
\newcommand{\vecN}{\mathbf{N}}
\newcommand{\vecP}{\mathbf{P}}
\newcommand{\vecQ}{\mathbf{Q}}
\newcommand{\vecS}{\mathbf{S}}
\newcommand{\vecT}{\mathbf{T}}
\newcommand{\vecW}{\mathbf{W}}
\newcommand{\vecZ}{\mathbf{Z}}
\newcommand{\vecb}{\mathbf{b}}
\newcommand{\vecc}{\mathbf{c}}
\newcommand{\vece}{\mathbf{e}}
\newcommand{\vecf}{\mathbf{f}}
\newcommand{\vecg}{\mathbf{g}}
\newcommand{\vech}{\mathbf{h}}
\newcommand{\veci}{\mathbf{i}}
\newcommand{\vecr}{\mathbf{r}}
\newcommand{\vecu}{\mathbf{u}}
\newcommand{\vecv}{\mathbf{v}}
\newcommand{\vecw}{\mathbf{w}}
\newcommand{\vecx}{\mathbf{x}}
\newcommand{\vecy}{\mathbf{y}}
\newcommand{\vecz}{\mathbf{z}}
\newcommand{\Ach}{\mathcal{H}}
\newcommand{\Ell}{\mathcal{L}}
\newcommand{\En}{\mathcal{N}}
\newcommand{\Pe}{\mathcal{P}}
\newcommand{\Ro}{\mathcal{R}_0}
\newcommand{\Rc}{\mathcal{R}_c}
\newcommand{\Rr}{\mathcal{R}}
\newcommand{\Qu}{\mathcal{Q}}
\newcommand{\Te}{\mathcal{T}}
\newcommand{\Ve}{\mathcal{V}}
\newcommand{\Dub}{\mathcal{W}}
\newcommand{\Zed}{\mathcal{Z}}
\newcommand{\Exp}{\mathbb{E}}
\newcommand{\de}{\mbox{d}}
\newcommand{\wrt}{\,\mbox{d}}
\newcommand{\veczero}{\mathbf{0}}
\newcommand{\vecone}{\mathbf{1}}
\newcommand{\vecalpha}{\mbox{\boldmath$\alpha$}}
\newcommand{\vectheta}{\mbox{\boldmath$\theta$}}
\newcommand{\vecphi}{\mbox{\boldmath$\phi$}}
\newcommand{\vecPhi}{\mbox{\boldmath$\Phi$}}
\newcommand{\vecSigma}{\mbox{\boldmath$\Sigma$}}
\newcommand{\vecimath}{\mbox{\boldmath$\imath$}}
\newcommand{\cosech}{\mbox{cosech}\,}
\newcommand{\sign}{\mbox{sign}\,}
\newcommand{\trace}{\mbox{trace}\,}
\newcommand{\Inf}{\mbox{Inf}\,}

@@

New Zealand has been working towards elimination of endemic (domestic) measles virus transmission, but has suffered from small, yet significant outbreaks of measles after measles introductions from abroad. In this draft final report we review the draft \emph {Progress Towards Measles Elimination in New Zealand - Final} report from the New Zealand Ministry of Health to the World Health Organization (WHO) Western Pacific Region. We identified additional analyses that may help understand risk of infection in New Zealand, and present the results of statistical analyses of risk factors for measles cases in New Zealand during outbreaks since 2007. We provide cost analyses for the measles outbreaks in New Zealand, and include modelling of measles outbreaks, including pre- and post-vaccination scenarios, based on the numbers of naive people at the District Health Board (DHB) and national level. We provide benefit--cost analyses using the results from those model simulations, along with a number of alternative vaccination strategies to achieve different vaccination coverage levels. Our key findings were:

\begin{itemize}
\item The \emph {Progress Towards Measles Elimination in New Zealand - Final} report was of high quality and contained substantial information and useful analyses.
\item Age is the best predictor of risk of measles infection in multivariate regression analyses, though some groups, such as people of Pacific ethnicity, the less socially deprived, and European and Maori school age children have been more likely to be measles cases in outbreaks since 2007.
\item New Zealand is at risk of frequent measles importation due to travel and endemic measles elsewhere in the globe.
\item Peak overall travel rates, and thus presumably risk from measles importation, is in December. However, New Zealander and immigrant or non--New Zealander travelling is otherwise out of phase, with peak travel for New Zealanders during the winter, and non--New Zealanders summer.
\item Analyses of outbreak data suggest that measles basic reproductive number ($R_v$, the number of secondary infections) values often include 1 and this year, 2014, as analysed from data from April, is well above one. This analysis suggests improved vaccination is a requisite to prevent measles becoming endemic again.
\item Estimates of the proportion of the currently naive New Zealand population requiring additional vaccination to ensure measles does not persist is approximately 28\%, leading to an additional 131,500 vaccinations.
\item The cost of the current 2013--2014 measles outbreak is estimated to be at least \$809,000.
\item The benefit--cost ratio analyses suggest additional vaccination is extremely beneficial financially.
\end{itemize}

@@

We agree with the report conclusions that testing for measles is performed appropriately within the required timeframe. Clearly improving inter-laboratory communication and collaboration and timeliness of the testing and reporting is necessary for rapid responses to measles introductions following measles control. Vaccination coverage presented in the report and to ourselves confirms that immunisation levels are approaching 94\% for MMR dose one (birth cohorts 2009 and 2010) and 89\% for MMR dose two (birth cohorts 2006 and 2007). However, only Asian and Pacific ethnicities have consistently had MMR dose one coverage approaching or exceeding 95\% for cohorts from 2007 onwards, and thus we agree with the report's conclusions that timeliness and coverage of vaccination need improving. This is particularly in light of our preliminary modelling and risk analyses results (section~\ref{sub:risk_analyses} and section~\ref{sec:epidemic_modelling}).

@@

We include multivariate modelling to account for confounding within the univariate analyses for measles cases in New Zealand, and descriptive analyses of risk of infection due to previous vaccination history, current rates of immunity within the population. In light of the apparent increasing trend in measles incidence in the last few years (Figure~\ref{fig:incidence1997}), we reviewed the information on measles importation and the origins of the introductions of measles into New Zealand. To help understand the risk of measles importation, with a particular goal of enabling the Ministry to better inform travellers and understand high risk periods, we sought to quantitatively evaluate the risk of measles importation from travel.

@@

%%%%%%%%%%%%%%%%%%
\subsection{Risk of measles infection in New Zealand analyses methods}
We received the raw EpiSurv measles case data from The Institute of Environmental Science and Research Ltd (ESR) on 27 June 2014. Initial analyses of those ESR data (not shown) suggested that denominator data were required to perform multivariate analyses to avoid confounding results due to a lack of independence among risk factors. Specifically we required $\textsf{Age} \times \textsf{Prioritised Ethnicity} \times \textsf{NZDep}$ data for New Zealand to test whether interactions among case covariates provide additional information on risk over the univariate analyses performed in the \emph{Progress Towards Measles Elimination in New Zealand - Final} report. These $\textsf{Age} \times \textsf{Prioritised Ethnicity} \times \textsf{NZDep}$ data were provided to us on 3 July 2014 by the University of Otago. We used these denominator data to determine if there were interactions among specific age categories, Prioritized Ethnicities, and NZDep that might exist among cases allowing better understanding of risk of measles infection.

@@

\begin{figure}
\begin{center}
<<fig=TRUE,echo=FALSE,width = 15>>=
par(mfrow=c(1,1))
par(mar=c(5,6,4,2)+0.1)
par(cex.axis=2)
hist(time$AgeInYears,col="grey",xlab="Age in years",main="",breaks=90,include.lowest=TRUE,right=F,ylab="",cex.lab=2)
hist(test$AgeInYears,col="black",breaks=90,include.lowest=TRUE,right=F,add=T)
legend("topright",c("1997-2014","2007-2014"),col=c("grey","black"),pch=15,bty="n",cex=1.8)
@
\end{center}
\caption{Age of measles cases in years in New Zealand for two periods, 1997--2014 and 2007--2014}
\label{fig:ageinyears}
\end{figure}

@@

\begin{figure}[h!]
\begin{center}
<<fig=TRUE,echo=FALSE,width = 11,height=9>>=
par(mar=c(5,6,4,2)+0.1)
par(cex.axis=2)
plot(tp$cases/tp$Popn,main="",ylab="Per Capita Rate",pch=16,cex=2,col=c(rep(1,10),rep(2,10),rep(3,10),rep(4,10),rep(6,10)),
     xlab="Category",cex.lab=2)
legend("topleft",c("Asian","European","Maori","MLA","Pacific"),
       col=c(1:4,6),pch=16,bty="n",cex=1.8)
@
\end{center}
\caption{Per capita rates of measles infections broken down by ethnicity, see Table~\ref{table:percap} for details}
\label{fig:percap}
\end{figure}

@@

@@

For our measles importation risk analyses, we use arrivals data from New Zealand immigration and New Zealander travel data by country and year (\href{http://www.immigration.govt.nz/}{www.immigration.govt.nz}) to measure human movement to and from New Zealand. We collated country population size, measles incidence and measles vaccination cover from the WHO (\href{http://www.who.int/research/en/}{www.who.int/research/en/}). Note the immigration figures use all immigration of foreign nationals, coming for whatever purpose and includes non-New Zealanders resident in New Zealand, but not yet holders of New Zealand passports. We used the WHO data to determine per capita measles cases for each year and used these data and the number of immigrants to New Zealand to begin to understand where measles was likely to be imported from. We used simple per capita rates for measles and the number of travellers to each country to score and map the risk of mealses importation. We use the data from 2012 because this year had the most complete WHO measles data and yet was most recent, thus accounts for improved measles vaccination coverage following the United Nations Millenium Development Goals' improvements in measles vaccination coverage.

@@

\begin{figure}[h!]
\begin{center}
<<fig=TRUE,echo=FALSE,width = 11>>=
print(np2)
@
\end{center}
\caption{Measles incidence per million 2012}
\label{fig:incidence12}
\end{figure}

>>>>

\usepackage{times}
\usepackage{hyperref}
\usepackage{tikz}
%\usepackage{bibunits}
\usepackage{natbib}
\usepackage{graphics}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{indentfirst}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{eurosym}
\usepackage{todonotes}
\usepackage{pdflscape}
\usepackage{booktabs}
\usepackage{array}
\usepackage{rotating}
\usepackage{threeparttable}
\usepackage{blindtext}
\usepackage{dcolumn}
\usepackage{tabularx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{caption2}
\usepackage{caption}
\usepackage{fancyhdr}
\usepackage{placeins}
\usepackage{framed}
\usepackage{float}
\newcommand{\vecA}{\mathbf{A}}
\newcommand{\vecB}{\mathbf{B}}
\newcommand{\vecC}{\mathbf{C}}
\newcommand{\vecD}{\mathbf{D}}
\newcommand{\vecE}{\mathbf{E}}
\newcommand{\vecF}{\mathbf{F}}
\newcommand{\vecH}{\mathbf{H}}
\newcommand{\vecI}{\mathbf{I}}
\newcommand{\vecJ}{\mathbf{J}}
\newcommand{\vecK}{\mathbf{K}}
\newcommand{\vecM}{\mathbf{M}}
\newcommand{\vecN}{\mathbf{N}}
\newcommand{\vecP}{\mathbf{P}}
\newcommand{\vecQ}{\mathbf{Q}}
\newcommand{\vecS}{\mathbf{S}}
\newcommand{\vecT}{\mathbf{T}}
\newcommand{\vecW}{\mathbf{W}}
\newcommand{\vecZ}{\mathbf{Z}}
\newcommand{\vecb}{\mathbf{b}}
\newcommand{\vecc}{\mathbf{c}}
\newcommand{\vece}{\mathbf{e}}
\newcommand{\vecf}{\mathbf{f}}
\newcommand{\vecg}{\mathbf{g}}
\newcommand{\vech}{\mathbf{h}}
\newcommand{\veci}{\mathbf{i}}
\newcommand{\vecr}{\mathbf{r}}
\newcommand{\vecu}{\mathbf{u}}
\newcommand{\vecv}{\mathbf{v}}
\newcommand{\vecw}{\mathbf{w}}
\newcommand{\vecx}{\mathbf{x}}
\newcommand{\vecy}{\mathbf{y}}
\newcommand{\vecz}{\mathbf{z}}
\newcommand{\Ach}{\mathcal{H}}
\newcommand{\Ell}{\mathcal{L}}
\newcommand{\En}{\mathcal{N}}
\newcommand{\Pe}{\mathcal{P}}
\newcommand{\Ro}{\mathcal{R}_0}
\newcommand{\Rc}{\mathcal{R}_c}
\newcommand{\Rr}{\mathcal{R}}
\newcommand{\Qu}{\mathcal{Q}}
\newcommand{\Te}{\mathcal{T}}
\newcommand{\Ve}{\mathcal{V}}
\newcommand{\Dub}{\mathcal{W}}
\newcommand{\Zed}{\mathcal{Z}}
\newcommand{\Exp}{\mathbb{E}}
\newcommand{\de}{\mbox{d}}
\newcommand{\wrt}{\,\mbox{d}}
\newcommand{\veczero}{\mathbf{0}}
\newcommand{\vecone}{\mathbf{1}}
\newcommand{\vecalpha}{\mbox{\boldmath$\alpha$}}
\newcommand{\vectheta}{\mbox{\boldmath$\theta$}}
\newcommand{\vecphi}{\mbox{\boldmath$\phi$}}
\newcommand{\vecPhi}{\mbox{\boldmath$\Phi$}}
\newcommand{\vecSigma}{\mbox{\boldmath$\Sigma$}}
\newcommand{\vecimath}{\mbox{\boldmath$\imath$}}
\newcommand{\cosech}{\mbox{cosech}\,}
\newcommand{\sign}{\mbox{sign}\,}
\newcommand{\trace}{\mbox{trace}\,}
\newcommand{\Inf}{\mbox{Inf}\,}

@@

New Zealand has been working towards elimination of endemic (domestic) measles virus transmission, but has suffered from small, yet significant outbreaks of measles after measles introductions from abroad. In this draft final report we review the draft \emph {Progress Towards Measles Elimination in New Zealand - Final} report from the New Zealand Ministry of Health to the World Health Organization (WHO) Western Pacific Region. We identified additional analyses that may help understand risk of infection in New Zealand. Here we present the results of statistical analyses of risk factors for measles cases in New Zealand during outbreaks since 2007. We provide cost analyses for the measles outbreaks in New Zealand, and include modelling of measles outbreaks, including pre- and post-vaccination scenarios, based on the numbers of naive people at the District Health Board (DHB) and national level. We provide benefit--cost analyses using the results from those model simulations, along with a number of alternative vaccination strategies to achieve different vaccination coverage levels. Our key findings were:

\begin{itemize}
\item The \emph {Progress Towards Measles Elimination in New Zealand - Final} report was of high quality and contained substantial information and useful analyses.
\item Age is the best predictor of risk of measles infection in multivariate regression analyses, though some groups, such as people of Pacific ethnicity, the less socially deprived, and European and Maori school age children have been more likely to be measles cases in outbreaks since 2007.
\item New Zealand is at risk of frequent measles importation due to travel and endemic measles elsewhere in the globe.
\item Peak overall travel rates, and thus presumably risk from measles importation, is in December. However, New Zealander and immigrant or non--New Zealander travelling is otherwise out of phase, with peak travel for New Zealanders during the winter, and non--New Zealanders summer.
\item Analyses of outbreak data suggest that measles basic reproductive number ($R_v$, the number of secondary infections) values often include 1 and this year, 2014, as analysed from data from April, is well above one. This analysis suggests improved vaccination is a requisite to prevent measles persisting in the population.
\item Estimates of the proportion of the currently naive New Zealand population requiring additional vaccination to ensure measles does not persist is approximately 28\%, leading to an additional 131,500 vaccinations.
\item The cost of the current 2013--2014 measles outbreak is estimated to be at least \$809,000.
\item The mean wage losses per measles case is estimated to be \$852
\item The mean cost per measles case attending hospital is estimated to be \$1710, and approximately 17\% of measles cases attend hospital
\item The mean management costs per case were \$1,765
\item The benefit--cost (B/C) ratio analyses suggest additional vaccination is extremely beneficial financially (B/C $>>$1).
\item Large outbreaks, with a mean size of approximately around cases per year, median of 2 cases, but peak size of up to many thousands, may occur regularly due to importation, despite $R_v$ being below one and the epidemic predicted to die out with additional vaccinations.
\end{itemize}

@@

We agree with the report conclusions that testing for measles is performed appropriately within the required timeframe. Clearly improving inter-laboratory communication and collaboration and timeliness of the testing and reporting is necessary for rapid responses to measles introductions following measles control. Vaccination coverage presented in the report and to ourselves confirms that immunisation levels are approaching 94\% for MMR dose one (birth cohorts 2009 and 2010) and 89\% for MMR dose two (birth cohorts 2006 and 2007). However, only Asian and Pacific ethnicities have consistently had MMR dose one coverage approaching or exceeding 95\% for cohorts from 2007 onwards, and thus we agree with the report's conclusions that timeliness and coverage of vaccination need improving. This is particularly in light of our modelling and risk analyses results (section~\ref{sub:risk_analyses} and section~\ref{sec:epidemic_modelling}).

@@

We include multivariate modelling to account for confounding within the univariate analyses for measles cases in New Zealand (section~\ref{sub:risk_analyses}), descriptive analyses of risk of infection due to previous vaccination history (section~\ref{sub:risk_analyses}), and current rates of immunity within the population (section~\ref{sub:risk_analyses}). In light of the apparent increasing trend in measles incidence in the last few years (Figure~\ref{fig:incidence1997}), we reviewed the information on measles importation and the origins of the introductions of measles into New Zealand. To help understand the risk of measles importation, with a particular goal of enabling the Ministry to better inform travellers and understand high risk periods, we sought to quantitatively evaluate the risk of measles importation from travel (section~\ref{sub:risk_analyses}).

@@

%%%%%%%%%%%%%%%%%%
\subsection{Risk of measles infection in New Zealand analyses methods}
\label{sub:name}

We received the raw EpiSurv measles case data from The Institute of Environmental Science and Research Ltd (ESR) on 27 June 2014. Initial analyses of those ESR data (not shown) suggested that denominator data were required to perform multivariate analyses to avoid confounding results due to a lack of independence among risk factors. Specifically we required $\textsf{Age} \times \textsf{Prioritised Ethnicity} \times \textsf{NZDep}$ data for New Zealand to test whether interactions among case covariates provide additional information on risk over the univariate analyses performed in the \emph{Progress Towards Measles Elimination in New Zealand - Final} report. These $\textsf{Age} \times \textsf{Prioritised Ethnicity} \times \textsf{NZDep}$ data were provided to us on 3 July 2014 by the University of Otago. We used these denominator data to determine if there were interactions among specific age categories, Prioritized Ethnicities, and NZDep that might exist among cases allowing better understanding of risk of measles infection.

@@

\begin{figure}
\begin{center}
<<fig=TRUE,echo=FALSE,width = 15,height=12>>=
par(mfrow=c(1,1))
par(mar=c(5,6,4,2)+0.1)
par(cex.axis=2)
hist(time$AgeInYears,col="grey",xlab="Age in years",main="",breaks=90,include.lowest=TRUE,right=F,ylab="",cex.lab=2)
hist(test$AgeInYears,col="black",breaks=90,include.lowest=TRUE,right=F,add=T)
legend("topright",c("1997-2014","2007-2014"),col=c("grey","black"),pch=15,bty="n",cex=1.8)
@
\end{center}
\caption{Age of measles cases in years in New Zealand for two periods, 1997--2014 and 2007--2014}
\label{fig:ageinyears}
\end{figure}

@@

\begin{figure}
\begin{center}
<<fig=TRUE,echo=FALSE,width = 11,height=9>>=
par(mar=c(5,6,4,2)+0.1)
par(cex.axis=2)
plot(tp$cases/tp$Popn,main="",ylab="Per Capita Rate",pch=16,cex=2,col=c(rep(1,10),rep(2,10),rep(3,10),rep(4,10),rep(6,10)),
     xlab="Category",cex.lab=2)
legend("topleft",c("Asian","European","Maori","MLA","Pacific"),
       col=c(1:4,6),pch=16,bty="n",cex=1.8)
@
\end{center}
\caption{Per capita rates of measles infections broken down by ethnicity, see Table~\ref{table:percap} for details}
\label{fig:percap}
\end{figure}

@@

\subsection{Vaccination history and measles infection}

In this section we describe the vaccination history of the measles cases from 2007--2014.

\subsection{Population immunity}

In this section we describe the population immunity levels at both a national level and that for each DHB in New Zealand. To do this we use the census data from NZ statistics (Statistics New Zealand, 2013) and the vaccination and serosurvey data provided by the Ministry of Health at the commencement of this work. Specifically, we use the data in Table \ref{table:sero}, that uses the serosurvey results and the National Immunisation Register for under 6 year olds.

\begin{table}[htdp]
\begin{center}
\begin{tabular}{lrrrr}
\hline
Age &  Proportion immune\\
\hline
0  & 0\\
1  & 0.89\\
2	& 0.92\\
3	& 0.93\\
4	& 0.93\\
5	& 0.92\\
6-13	& 0.8\\
14-18	& 0.83\\
19-23	& 0.77\\
24-32	& 0.85\\
33-52	& 0.92\\
>52	& 0.99\\
\hline
\end{tabular}
\end{center}
\caption{Vaccination coverage and serosurvey estimates of immunity among different age classes}
\label{table:sero}
\end{table}%

@@

For our measles importation risk analyses, we use arrivals data from New Zealand immigration and New Zealander travel data by country and year to measure human movement to and from New Zealand  (\href{http://www.immigration.govt.nz/}{www.immigration.govt.nz}). We collated country population size, measles incidence and measles vaccination cover from the WHO (\href{http://www.who.int/research/en/}{www.who.int/research/en/}). Note the immigration figures use all immigration of foreign nationals, coming for whatever purpose, and includes non-New Zealanders resident in New Zealand, but not yet holders of New Zealand passports. We used the WHO data to determine per capita measles cases for each year and used these data and the number of immigrants to New Zealand to begin to understand where measles was likely to be imported from. We used simple per capita rates for measles and the number of travellers to each country to score and map the risk of mealses importation. We use the data from 2012 because this year had the most complete WHO measles data and yet was most recent, thus accounts for improved measles vaccination coverage following the United Nations Millenium Development Goals' improvements in measles vaccination coverage.

The distribution of the cases per category used in the regression analyses are in Figure~\ref{fig:histcasecat}.
\begin{figure}
\begin{center}
<<fig=TRUE,echo=FALSE,width = 11>>=
par(cex.axis=2)
par(mar=c(5,6,4,2)+0.1)
hist(tpsub$cases,xlab="Cases",main='',col='grey',breaks=20,ylab="",cex.lab=2)
@
\end{center}
\caption{Distribution of cases per category used in the final regression model}
\label{fig:histcasecat}
\end{figure}

The predicted values from the regression model plotted against the reported cases are shown in Figure~\ref{fig:predict}, and the residuals are shown in Figure~\ref{fig:resid}.
\begin{figure}
\begin{center}
<<fig=TRUE,echo=FALSE,width =11, height=11>>=
res<-predict(model3)
par(mar=c(5,6,4,2)+0.1)
par(cex.axis=2)
plot(exp(res),tpsub$cases,xlab="cases",ylab='predictions',main="",pch=16,col="darkgrey",cex.lab=2,cex=2)
@
\end{center}
\caption{Regression model (Equation~\ref{eq:reg}) predictions plotted against the cases (Table~\ref{table:percap})}
\label{fig:predict}
\end{figure}

\begin{figure}
\begin{center}
<<fig=TRUE,echo=FALSE,width = 11>>=
par(mar=c(5,6,4,2)+0.1)
par(cex.axis=2)
hist(model3$residuals,main="",xlab="residuals",col="grey",cex.lab=2,ylab="")
@
\end{center}
\caption{Histogram of residuals from the regression model (Equation~\ref{eq:reg})}
\label{fig:resid}
\end{figure}

The significance of the different predictor variables can be seen in the ANOVA results:
\par
\vspace{5mm} %5mm vertical space
<<label=regmodelanova,include=FALSE,echo=FALSE,results=tex>>=
library(xtable)
resan<-xtable(anova(model3,test="F"))
print(resan,floating = FALSE)
@
\vspace{5mm} %5mm vertical space
A summary of the regression model (Equation~\ref{eq:reg}) with the individual effects and the statistical support for the estimated coefficients can be seen here:
\vspace{5mm} %5mm vertical space
<<label=regmodel,include=FALSE,results=tex,echo=FALSE>>=
library(xtable)
xtb<-xtable(summary(model3))
print(xtb,floating = FALSE)
@
\vspace{5mm} %5mm vertical space

\par Apart from over-representation of some MLA categories discussed above and not included here, the results of the regression model suggest that age is a strong predictor of being a measles case. Indeed, all age categories are significantly less likely to be measles cases compared to 0-2 year olds, and the likelihood decreases with age (Figure~\ref{fig:ageinyears}).

\subsection{Vaccination history and measles infection results}

The majority of the measles cases (82.8\%, 955/1154) from 2007--2014 were in unvaccinated people (Figure~\ref{fig:ageandvac}). However, 12.6\% (154/1154) cases had received their first dose of vaccine and 4.7\% had received their second. A further breakdown of those data suggest that majority of those 'vaccine failures' were vaccinated around the first year of age (Figure~\ref{fig:vaccstat}).

\begin{figure}
\begin{center}
<<fig=TRUE,echo=FALSE,width = 11>>=
AgeVac<-table(testv$VC,testv$AgeInYears)
row.names(AgeVac)<-c("Unvaccinated","Dose1","Dose2")
AgeVac<-t(AgeVac)
par(mar=c(5,6,4,2)+0.1)
par(cex.axis=2)
barplot(t(AgeVac),xlab="Age",col=c("darkgrey","red","orange"),main="",cex.lab=2)
legend("topright",c("Unvaccinated","Dose 1","Dose 2"),fill=c("darkgrey","red","orange"),cex=1.8,bty="n")
@
\end{center}
\caption{Age and vaccination status of cases}
\label{fig:ageandvac}
\end{figure}

\begin{figure}
\begin{center}
<<fig=TRUE,echo=FALSE,width = 11>>=
par(mar=c(5,6,4,2)+0.1)
par(cex.axis=2)
hist(vac$Dose1Mths,breaks=100,col=rgb(0,0,1,1/4),xlab="Age in months",main="",cex.lab=2,ylab="")
abline(v=12,col="blue")
hist(vac$Dose2Mths,breaks=50,add=T,col=rgb(1,0,0,1/4))
abline(v=60,col="red")
legend("topright",c("Dose 1","Dose 2"),col=c(rgb(0,0,1,1/4),rgb(1,0,0,1/4)),pch=15,bty="n",cex=1.8)
legend("top",c("12 months","60 months"),col=c("blue","red"),lty=1,bty="n",cex=1.8)
@
\end{center}
\caption{Age of vaccination of vaccinated measles cases}
\label{fig:vaccstat}
\end{figure}

\subsection{Population immunity results}

The majority of the naive among the general New Zealand population are in their first years of life (Figure~\ref{fig:naive}). However, the distribution of naive at a national level shows that the recent MMR vaccination schemes are reducing the proportions of naive in the very young to greater levels than in older young people  (Figure~\ref{fig:naive}). Plotting the breakdown of these figures by DHB clearly shows that the greatest numbers of naive people are in DHBs with larger urban areas (Figure~\ref{fig:naive_map}). The distribution of the numbers of naive and the total naive populations per DHB are given in \ref{fig:Northland}\ref{fig:Waitemata}\ref{fig:Auckland}\ref{fig:Counties_Manukau}\ref{fig:Waikato}\ref{fig:Lakes}\ref{fig:BayofPlenty}\ref{fig:Tairawhiti}\ref{fig:Taranaki}\ref{fig:HawkesBay}\ref{fig:Whanganui}\ref{fig:Midcentral}\ref{fig:Hutt}\ref{fig:CapitalandCoast}\ref{fig:Wairarapa}\ref{fig:NelsonMarlborough}\ref{fig:WestCoast}\ref{fig:Canterbury}\ref{fig:SouthCanterbury}\ref{fig:Southern}.

\begin{figure}
\begin{center}
<<fig=TRUE,echo=FALSE,width = 11>>=
par(cex.axis=2)
par(mar=c(5,6,4,2)+0.1)
plot(pop,xlab="Age",ylab="Population",cex.lab=2)
points(naive,pch=16)
legend("topright",c("Population","Naive"),pch=c(1,16),bty="n",cex=1.8)
@
\end{center}
\caption{New Zealand population by age and estimated numbers of naive people in each age class}
\label{fig:naive}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=1.2\textwidth]{naive_map.pdf}
     \end{center}
     \caption{Numbers of naive individuals per District Health Board}
     \label{fig:naive_map}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb1.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Northland District Health Board}
     \label{fig:Northland}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb2.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Waitemata District Health Board}
     \label{fig:Waitemata}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb3.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Auckland District Health Board}
     \label{fig:Auckland}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb4.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Counties Manukau District Health Board}
     \label{fig:Counties_Manukau}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb5.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Waikato District Health Board}
     \label{fig:Waikato}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb6.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Lakes District Health Board}
     \label{fig:Lakes}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb7.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Bay of Plenty District Health Board}
     \label{fig:BayofPlenty}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb8.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Tairawhiti District Health Board}
     \label{fig:Tairawhiti}
\end{figure}

\begin{figure}[H]
  \begin{center}
    \includegraphics[width=0.6\textwidth]{dhb9.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Taranaki District Health Board}
     \label{fig:Taranaki}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb10.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Hawke's Bay District Health Board}
     \label{fig:HawkesBay}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb11.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Whanganui District Health Board}
     \label{fig:Whanganui}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb12.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Midcentral District Health Board}
     \label{fig:Midcentral}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb13.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Hutt District Health Board}
     \label{fig:Hutt}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb14.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Capital and Coast District Health Board}
     \label{fig:CapitalandCoast}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb15.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Wairarapa District Health Board}
     \label{fig:Wairarapa}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb16.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Nelson Marlborough District Health Board}
     \label{fig:NelsonMarlborough}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb17.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, West Coast District Health Board}
     \label{fig:WestCoast}
\end{figure}

\begin{figure}[H]
   \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb18.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Canterbury District Health Board}
     \label{fig:Canterbury}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb19.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, South Canterbury District Health Board}
     \label{fig:SouthCanterbury}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.6\textwidth]{dhb20.pdf}
     \end{center}
     \caption{Numbers of naive individuals per age class, Southern District Health Board}
     \label{fig:Southern}
\end{figure}

\subsection{Measles importation risk results}

@@

\begin{figure}
\begin{center}
<<fig=TRUE,echo=FALSE,width = 11>>=
print(np2)
@
\end{center}
\caption{Measles incidence per million 2012}
\label{fig:incidence12}
\end{figure}

\begin{figure}
\begin{center}
<<fig=TRUE,echo=FALSE,width = 11>>=
print(np3)
@
\end{center}
\caption{Measles vaccination cover (\%) 2012 }
\label{fig:cover12}
\end{figure}
